---
title: "Mean Losses per Peril"
author: "Mirco Heidemann"
date: "4/25/2022"
output: pdf_document
---
Mean losses per peril:
- Flood, storm and hail
- With zero claims and without

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

Read **loss** into data frame, read gvz index.
```{r read}
file_name_loss <- paste0("Data/", "20220425_GemDat_Schaden_Export_Ng.csv")
index_gvz <- paste0("Data/", "Versicherungsindex_GVZ.csv")

# Read loss data
df_orig_loss <- read_delim(file_name_loss, delim = ";")

# Read "GVZ Versicherungs Index"
df_index <- read_delim(index_gvz, delim = ";",
                       col_select = c(1,3)) %>% 
  mutate(Jahr = format(as.Date(Jahr, "%d.%m.%Y"), "%Y")) %>% 
  rename(Index = `Versicherungsindex GVZ`)
```

Filter natural hazards between 1981 and 2021, index claims to 2021.
Data Frames of all claims and from the last 10 years only.
```{r}
# All losses
df_loss <- df_orig_loss %>% 
  mutate(ClaimIncidentDate = as.Date(ClaimIncidentDate, "%d.%m.%Y"),
         ClaimIncidentYear = format(ClaimIncidentDate, "%Y")) %>% 
  filter(ClaimCausation %in% c("Sturm", "Hagel", "Überschwemmung, Hochwasser"),
           ClaimIncidentDate > as.Date("31.12.1981", "%d.%m.%Y") & 
           ClaimIncidentDate < as.Date("01.01.2022", "%d.%m.%Y"),
         ClaimSumBuilding >= 0,
         !is.na(InsuranceValue)) %>% 
  left_join(df_index, by = c("ClaimIncidentYear" = "Jahr")) %>% 
  mutate(schad_index = round(max(Index) / Index * ClaimSumBuilding),
         schadengrad = ifelse(InsuranceValue > 0, schad_index/InsuranceValue, 0))

# Nat hazard claims without zero losses and without zeros insurance values
df_loss_nonzero <- df_loss %>% 
  filter(ClaimSumBuilding > 0,
         InsuranceValue > 0)

# All losses from 2012 to 2021
df_loss_last10 <- df_loss %>% 
  filter(ClaimIncidentDate > as.Date("31.12.2011", "%d.%m.%Y") &
           ClaimIncidentDate < as.Date("01.01.2022", "%d.%m.%Y"))

# Nat hazard claims without zero losses and without zeros insurance values from 2012 to 2021
df_loss_nonzero_last10 <- df_loss_nonzero %>% 
  filter(ClaimIncidentDate > as.Date("31.12.2011", "%d.%m.%Y") &
           ClaimIncidentDate < as.Date("01.01.2022", "%d.%m.%Y"))
```

Summary of the data frames
```{r}
# summary(df_loss)
# summary(df_loss_nonzero)
# summary(df_loss_last10)
# summary(df_loss_nonzero_last10)
```

Mean, median and mean damage degree of losses with zeros and without zeros. Once over all years and once over the last 10 year only.
```{r}
tbl_loss <- df_loss %>% 
  group_by(ClaimCausation) %>% 
  summarize(Schadenmeldungen = n(),
            Mean_Schaden = round(mean(schad_index)),
            Median_Schaden = round(median(schad_index)),
            MDD_pm = round(mean(schadengrad) * 1000, 2))

tbl_loss_nonzero <-  df_loss_nonzero %>% 
  group_by(ClaimCausation) %>% 
  summarize(Schadenmeldungen = n(),
            Mean_Schaden = round(mean(schad_index)),
            Median_Schaden = round(median(schad_index)),
            MDD_pm = round(mean(schadengrad) * 1000, 2))

tbl_loss_last10 <- df_loss_last10 %>% 
  group_by(ClaimCausation) %>% 
  summarize(Schadenmeldungen = n(),
            Mean_Schaden = round(mean(schad_index)),
            Median_Schaden = round(median(schad_index)),
            MDD_pm = round(mean(schadengrad) * 1000, 2))

tbl_loss_nonzero_last10 <- df_loss_nonzero_last10 %>% 
  group_by(ClaimCausation) %>% 
  summarize(Schadenmeldungen = n(),
            Mean_Schaden = round(mean(schad_index)),
            Median_Schaden = round(median(schad_index)),
            MDD_pm = round(mean(schadengrad) * 1000, 2))

```

