---
title: "Mean Losses per Peril"
author: "Mirco Heidemann"
date: "4/25/2022"
output: pdf_document
---
Mean losses per peril:
- Flood, storm and hail
- With zero claims and without

### Start Coding...
RMarkdown files are a special case, as they work slightly differently to .R files in terms of file paths, i.e. they behave like mini projects of their own, where the default working directory is where the Rmd file is saved. To save RMarkdown files in a basic folder structure set up, it’s recommended to use the [here package](https://github.com/jennybc/here_here) and its workflow.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
```

Read **loss** into data frame, read gvz index.
```{r read}
# Path workflow for RMarkdown files with here packages
file_name_loss <- here("Data", "20220425_GemDat_Schaden_Export_Ng.csv")
index_gvz <- here("Data", "Versicherungsindex_GVZ.csv")

# Read loss data
df_orig_loss <- read_delim(file_name_loss, delim = ";")

# Read "GVZ Versicherungs Index"
df_index <- read_delim(index_gvz, delim = ";",
                       col_select = c(1,3)) %>% 
  mutate(Jahr = format(as.Date(Jahr, "%d.%m.%Y"), "%Y")) %>% 
  rename(Index = `Versicherungsindex GVZ`)
```

Filter natural hazards between 1981 and 2021, index claims to 2021.
Data Frames of all claims and from the last 10 years only.
```{r}
# All losses
df_loss <- df_orig_loss %>% 
  mutate(ClaimIncidentDate = as.Date(ClaimIncidentDate, "%d.%m.%Y"),
         ClaimIncidentYear = format(ClaimIncidentDate, "%Y")) %>% 
  filter(ClaimCausation %in% c("Sturm", "Hagel", "Überschwemmung, Hochwasser"),
           ClaimIncidentDate > as.Date("31.12.1981", "%d.%m.%Y") & 
           ClaimIncidentDate < as.Date("01.01.2022", "%d.%m.%Y"),
         ClaimSumBuilding >= 0,
         !is.na(InsuranceValue)) %>% 
  left_join(df_index, by = c("ClaimIncidentYear" = "Jahr")) %>% 
  mutate(schad_index = round(max(Index) / Index * ClaimSumBuilding),
         schadengrad = ifelse(InsuranceValue > 0, schad_index/InsuranceValue, 0))

# Nat hazard claims without zero losses and without zeros insurance values
df_loss_nonzero <- df_loss %>% 
  filter(ClaimSumBuilding > 0,
         InsuranceValue > 0)

# All losses from 2012 to 2021
df_loss_last10 <- df_loss %>% 
  filter(ClaimIncidentDate > as.Date("31.12.2011", "%d.%m.%Y") &
           ClaimIncidentDate < as.Date("01.01.2022", "%d.%m.%Y"))

# Nat hazard claims without zero losses and without zeros insurance values from 2012 to 2021
df_loss_nonzero_last10 <- df_loss_nonzero %>% 
  filter(ClaimIncidentDate > as.Date("31.12.2011", "%d.%m.%Y") &
           ClaimIncidentDate < as.Date("01.01.2022", "%d.%m.%Y"))
```

Summary of the data frames
```{r}
# summary(df_loss)
# summary(df_loss_nonzero)
# summary(df_loss_last10)
# summary(df_loss_nonzero_last10)
```

Mean, median and mean damage degree of losses with zeros and without zeros. Once over all years and once over the last 10 year only.
```{r}
tbl_loss <- df_loss %>% 
  group_by(ClaimCausation) %>% 
  summarize(Schadenmeldungen = n(),
            Mean_Schaden = round(mean(schad_index)),
            Median_Schaden = round(median(schad_index)),
            MDD_pm = round(mean(schadengrad) * 1000, 2))

tbl_loss_nonzero <-  df_loss_nonzero %>% 
  group_by(ClaimCausation) %>% 
  summarize(Schadenmeldungen = n(),
            Mean_Schaden = round(mean(schad_index)),
            Median_Schaden = round(median(schad_index)),
            MDD_pm = round(mean(schadengrad) * 1000, 2))

tbl_loss_last10 <- df_loss_last10 %>% 
  group_by(ClaimCausation) %>% 
  summarize(Schadenmeldungen = n(),
            Mean_Schaden = round(mean(schad_index)),
            Median_Schaden = round(median(schad_index)),
            MDD_pm = round(mean(schadengrad) * 1000, 2))

tbl_loss_nonzero_last10 <- df_loss_nonzero_last10 %>% 
  group_by(ClaimCausation) %>% 
  summarize(Schadenmeldungen = n(),
            Mean_Schaden = round(mean(schad_index)),
            Median_Schaden = round(median(schad_index)),
            MDD_pm = round(mean(schadengrad) * 1000, 2))

```

